Magnetised Polish doughnuts revisited

### 2017-07-12 paper_accepted/paper_accepted.tex
boldface removed. Added referee acknowledgement

\newcommand{\tf}[1]{\textcolor{red}{TF: #1}}
\newcommand{\sg}[1]{\textcolor{blue}{SG: #1}}

@@

The equations of ideal general relativistic MHD are the following conservation laws, $\nabla_{\mu} T^{\mu\nu} = 0$, $\nabla_{\mu} \,^\ast F^{\mu\nu} = 0$, and 
$\nabla_{\mu} (\rho u^{\mu}) = 0$, 
where $\nabla_{\mu}$ is the covariant derivative and
\begin{equation}\label{eq:e-m_tensor}
T^{\mu\nu} = (w + b^2)u^{\mu}u^{\nu} + \left(p + \frac{1}{2}b^2\right)g^{\mu\nu} - b^{\mu}b^{\nu},
\end{equation}
is the energy-momentum tensor of a magnetised perfect fluid, $w$ and $p$ being the {\bf fluid enthalpy density} and fluid pressure, respectively. 
Moreover, $^\ast F^{\mu\nu} = b^{\mu}u^{\nu} - b^{\nu}u^{\mu}$ is the (dual of the) Faraday tensor relative to an observer with 
four-velocity $u^{\mu}$, and $b^{\mu}$ is the magnetic field in that frame, with
$b^2=b^{\mu}b_{\mu}$. Assuming the magnetic field is purely azimuthal, i.e.~$b^r = b^{\theta} = 0$,
and taking into account that the flow is stationary and axisymmetric, the conservation of the current density and of the Faraday tensor follow. Contracting Eq.~\eqref{eq:e-m_tensor} with the projection tensor $h^{\alpha}_{\,\,\beta} = \delta^{\alpha}_{\,\,\beta} + u^{\alpha}u_{\beta}$, we arrive at
\begin{equation}
(w + b^2)u_{\nu}\partial_i u^{\nu} + \partial_i\left(p + \frac{b^2}{2}\right) - b_{\nu}\partial_i b^{\nu}=0\,,
\end{equation}
where $i = r, \theta$. Following~\cite{Komissarov:2006} we rewrite this equation in terms of the specific angular momentum $l$ and of the angular velocity $\Omega$, to obtain
\begin{equation}\label{eq:diff_ver}
\partial_i(\ln u_t|) - \frac{\Omega \partial_i l}{1-l\Omega} + \frac{\partial_i p}{w} + \frac{\partial_i(\mathcal{L}b^2)}{2\mathcal{L}w} = 0\,,
\end{equation}
where $\mathcal{L} = g_{t\phi}^2 - g_{tt}g_{\phi\phi}$.
To integrate Eq.~\eqref{eq:diff_ver} we first assume a barotropic equation of state $w = w(p)$ of the form
\begin{equation}\label{eq:eos_fluid}
p = K w^{\kappa},
\end{equation}
with $K$ and $\kappa$ constants.
Then, we define the magnetic pressure as $p_{\mathrm{m}} = b^2/2$, and introduce the definitions $\tilde{p}_{\mathrm{m}} = \mathcal{L} p_{\mathrm{m}}$ and $\tilde{w} = \mathcal{L} w$, in order to write an analogue equation to Eq.~\eqref{eq:eos_fluid} for $\tilde{p}_{\mathrm{m}}$~\citep{Komissarov:2006}
\begin{equation}\label{eq:eos_mag_tilde}
\tilde{p}_{\mathrm{m}} = K_{\mathrm{m}} \tilde{w}_{\mathrm{m}}^{\lambda
},
\end{equation}
or, in terms of the magnetic pressure $p_{\mathrm{m}}$
\begin{equation}\label{eq:eos_mag}
p_{\mathrm{m}} = K_{\mathrm{m}} \mathcal{L}^{\lambda
-1} w^{\lambda
},
\end{equation}
where $K_{\mathrm{m}}$ and $\lambda
$ are constants.
This particular choice of barotropic relationships, $w = w(p)$ and $\tilde{w} = \tilde{w}(\tilde{p}_{\mathrm{m}})$, fulfill the general relativistic version of the von Zeipel theorem for a toroidal magnetic field~\citep{vonZeipel:1924, Zanotti:2015}, i.e.~the surfaces of constant $\Omega$ and constant $l$ coincide.

We can now integrate Eq.~\eqref{eq:diff_ver} to obtain
\begin{equation}\label{eq:pre_full_int}
\ln |u_t| - \int^l_0 \frac{\Omega \mathrm{d}l}{1 - \Omega l} + \int^p_0 \frac{\mathrm{d}p}{w} + \int_0^{\tilde{p}_{\mathrm{m}}} \frac{\mathrm{d}\tilde{p}_{\mathrm{m}}}{\tilde{w}} = \mathrm{const}.
\end{equation}
On the surface of the disc, and particularly on its inner edge, the conditions
$p = \tilde{p}_{\mathrm{m}} = 0, \; u_t = u_{t, \mathrm{in}}, \; l = l_{\mathrm{in}}$
are satisfied and, therefore, the integration constant is simply given by
\begin{equation}
\mathrm{const.} = \ln |u_t| - \int^l_{l_\mathrm{in}} \frac{\Omega \mathrm{d}l}{1 - \Omega l}\,.
\end{equation}
We can also introduce the total (gravitational plus centrifugal) potential $W$~\citep{Abramowicz:1978} and write the integral form of the equation of motion (the relativistic Euler equation) as
\begin{equation}\label{eq:potential}
W - W_{\mathrm{in}} = \ln|u_t| - \ln|u_{t,\mathrm{in}}| - \int^{l}_{l_{\mathrm{in}}} \frac{\Omega \mathrm{d}l}{1 - \Omega l},
\end{equation}
where $W_{\mathrm{in}}$ is the potential at the inner edge of the disc (in the equatorial plane). With this definition, we can write Eq.~\eqref{eq:pre_full_int} as
\begin{equation}\label{eq:full_int}
W - W_{\mathrm{in}} = \int^p_0 \frac{\mathrm{d}p}{w} + \int_0^{\tilde{p}_{\mathrm{m}}} \frac{\mathrm{d}\tilde{p}_{\mathrm{m}}}{\tilde{w}},
\end{equation}
which for a barotropic equation of state can be easily integrated to give
\begin{equation}\label{eq:pre_enthalpy_eq}
W - W_{\rm{in}} + \frac{\kappa}{\kappa - 1}\frac{p}{w} + \frac{\lambda
}{\lambda
 - 1}\frac{p_{\mathrm{m}}}{w} = 0\,.
\end{equation}
Replacing $p$ and $p_{\mathrm{m}}$ by equations \eqref{eq:eos_fluid} and \eqref{eq:eos_mag}, the previous equation reduces to
\begin{equation}\label{eq:enthalpy_eq}
W - W_{\rm{in}} + \frac{\kappa}{\kappa - 1} K w^{\kappa - 1} + \frac{\lambda
}{\lambda
 - 1}K_{\mathrm{m}}(\mathcal{L} w)^{\lambda
 - 1} = 0,
\end{equation}
which relates the distribution of the potential with the distribution of the {\bf enthalpy density}.

@@

\begin{figure*}[t]
\centering
\includegraphics[scale=0.14]{figures/fig1a.pdf}
\hspace{-0.3cm}
\includegraphics[scale=0.14]{figures/fig1b.pdf}
\hspace{-0.2cm}
\includegraphics[scale=0.14]{figures/fig1c.pdf}
\caption{Comparison with Komissarov's solution for a constant angular momentum model. The left panel shows the rest-mass density distribution in logarithm scale for all radii and all angles while the middle and right panels show, respectively, the angular profile of the logarithm of the rest-mass density at the centre of the disc and the radial profile of {\bf the logarithm of} the rest-mass density at the equatorial plane.}
\label{komissarov}
\end{figure*}

Next, we choose all the initial radial values for the integration of Eq.~\eqref{eq:F} to lie between $r_{\mathrm{cusp}}$ and $r_{\mathrm{c}}$ ($\theta = \pi / 2$). Since we are only interested in the equipotential surfaces inside the Roche lobe of the disc, our choice of initial values provides us a mapping of the equipotential surfaces of the torus. Given that we have already obtained both the equipotential surfaces $\theta(r)$ which cross the segment $[r_{\mathrm{cusp}}, r_{\mathrm{c}}]$ at the equatorial plane and the values of the potential in that segment, we can obtain the complete potential distribution for the torus (outside of that segment). Once we have the potential distribution, we can find the fluid pressure at the centre of the disc from Eq.~\eqref{eq:enthalpy_eq},
\begin{equation}
p_{\mathrm{c}} = w_{\mathrm{c}}(W_{\mathrm{in}} - W_{\mathrm{c}})\left(\frac{\kappa}{\kappa - 1} + \frac{\lambda
}{\lambda
 - 1}\frac{1}{\beta_{\mathrm{m}_{\mathrm{c}}}}\right)^{-1},
\end{equation}
where $w_{\mathrm{c}}$ is the {\bf enthalpy density} at the centre and
\begin{equation}
\label{eq:beta_eq}
\beta_{\mathrm{m}} = p/p_{\mathrm{m}},
\end{equation}
is the magnetisation parameter ($\beta_{\mathrm{m}_{\mathrm{c}}}$ being the magnetisation parameter at the centre of the disc). Using this definition, we can find the magnetic pressure at the centre,
\begin{equation}
p_{\mathrm{m_{\mathrm{c}}}} = p_{\mathrm{c}}/\beta_{\mathrm{m}_{\mathrm{c}}}\,.
\end{equation}
With both pressures known at the centre, fluid and magnetic, we can now find the constants $K$ and $K_{\mathrm{m}}$ using equations \eqref{eq:eos_fluid} and \eqref{eq:eos_mag}. Therefore, for a given inner radius of the disc $r_{\mathrm{in}}$ we can obtain the potential $W_{\mathrm{in}}$. We now finally have all the ingredients required to find the {\bf enthalpy density} distribution (Eq.~\eqref{eq:enthalpy_eq}), the fluid pressure and magnetic pressure distributions (Eqs.~\eqref{eq:eos_fluid} and \eqref{eq:eos_mag}), and the rest-mass density distribution, which can be trivially obtained inverting the barotropic relation, $p = K \rho^{\kappa}$. We note that the fluid {\bf enthalpy density} $w$ includes the rest-mass density $\rho$, and can thus be defined as $w=\rho h$ where $h$ is the specific enthalpy. The relativistic definition of this quantity is $h=1+\varepsilon+p/\rho$, where $\varepsilon$ is
the specific internal energy. From a thermodynamical point of view a non-relativistic fluid satisfies $h=1$. Therefore,
since we use poytropic equations of state (relating $p$ with either $w$ or $\rho$ in the same functional form) we are
implictly assuming that $h=1$, i.e.~the discs we build in our procedure are non-relativistic from a thermodynamical point of view.
%{\bf We note that the specific internal energy $\varepsilon$ is needed to compute the enthalpy. We obtain
%this quantity by combining the ideal fluid and the barotropic equations of state, which yields 
%$\varepsilon=\frac{K}{\kappa-1}\rho^{\kappa-1}$.}

@@

To reduce the number of free parameters to build the initial models we do as in~\citet{Komissarov:2006} and fix the values of the equation of state exponents, $\kappa$ and $\lambda$, and of the {\bf enthalpy density} at the disc centre, $w_{\mathrm{c}}$. More precisely, we choose $\kappa = \lambda = 4/3$ and $w_{\mathrm{c}} = 1$. This still leaves us with five parameters to control the size, shape, thickness, and magnetisation of the disc, namely the magnetisation parameter $\beta_{\mathrm{m}_{\mathrm{c}}}$, the parameters of the angular momentum distribution $\beta$ and $\gamma$, the black hole spin parameter $a$, and the inner radius of the disc $r_{\mathrm{in}}$. We build a series of 45 models, whose main features are summarised in Table~\ref{table:1} (for $\beta_{\mathrm{m}_{\mathrm{c}}} = 10^{3}$, i.e.~models where the effects of the magnetisation are unimportant), Table~\ref{table:2} ($\beta_{\mathrm{m}_{\mathrm{c}}} = 1$, i.e.~equipartition models), and Table~\ref{table:3} ($\beta_{\mathrm{m}_{\mathrm{c}}} = 10^{-3}$, i.e.~highly-magnetised models). {\bf We note that we can build models with very small values of $\beta_{\mathrm{m}_{\mathrm{c}}}$, such as  e.g.~$\beta_{\mathrm{m}_{\mathrm{c}}} = 10^{-20}$ (extremely high magnetisation) without encountering numerical difficulties. No qualitative differences in the structure of the discs are found once $\beta_{\mathrm{m}_{\mathrm{c}}}$ becomes smaller than a sufficiently small value, about $\beta_{\mathrm{m}_{\mathrm{c}}}=10^{-3}$. This is the reason why we choose that particular value as our lower limit in the figures and tables of the manuscript.}

@@

We start by assessing our procedure by first building an extra disc model that can be directly compared with one of the two models in~\citet{Komissarov:2006}. This is shown in Figure~\ref{komissarov}, which corresponds to the same constant specific angular momentum model A presented by~\citet{Komissarov:2006}. The parameters of this model are $a=0.9$, $\beta_{\mathrm{m}_{\mathrm{c}}}=0.1$, and $l=2.8$.
The visual comparison shows that our approach can reproduce those previous results with good agreement. The rest-mass density distribution in the $(r\sin\theta,r\cos\theta)$ plane shown in the left panel of Fig.~\ref{komissarov} is remarkably similar to that shown in the left panel of Fig.~2 in~\citet{Komissarov:2006}. Not only the morphology of both models is nearly identical but also the range of variation of the rest-mass density and the location of the disc centre agree well in both cases. This can be most clearly seen in the middle and right panels of Fig.~\ref{komissarov} which show, repectively, the angular profile at $r_{\rm c}$ and the radial profile at $\theta=\pi/2$. {\bf The middle figure} can be directly compared with Fig.~3 of~\citet{Komissarov:2006}. It is relevant to mention that very small changes in the location of the inner radius $r_{\mathrm{in}}$ have a significant effect on the maximum value of the rest-mass density, which explains the small differences between our figures and the ones presented in~\citet{Komissarov:2006}.

@@

% It must be noted that the location of the maximum of the magnetic pressure is nearly identical for all models considered when $\log_{10}\beta_{\mathrm{m}_{\mathrm{c}}} \sim 0.5$. At this value of $\beta_{\mathrm{m}_{\mathrm{c}}}$ all models seem to cross, as it is more clearly shown in the inset of the right panel of Fig.~\ref{max-vs-magnetisation}.

@@

\begin{acknowledgements}
It is a pleasure to thank P.~C.~Fragile and N.~Stergioulas for useful discussions and comments. Work supported by the Spanish MINECO (grant AYA2015-66899-C2-1-P) and the Generalitat Valenciana (PROMETEOII-2014-069).
\end{acknowledgements}

>>>>

@@

The equations of ideal general relativistic MHD are the following conservation laws, $\nabla_{\mu} T^{\mu\nu} = 0$, $\nabla_{\mu} \,^\ast F^{\mu\nu} = 0$, and 
$\nabla_{\mu} (\rho u^{\mu}) = 0$, 
where $\nabla_{\mu}$ is the covariant derivative and
\begin{equation}\label{eq:e-m_tensor}
T^{\mu\nu} = (w + b^2)u^{\mu}u^{\nu} + \left(p + \frac{1}{2}b^2\right)g^{\mu\nu} - b^{\mu}b^{\nu},
\end{equation}
is the energy-momentum tensor of a magnetised perfect fluid, $w$ and $p$ being the fluid enthalpy density and fluid pressure, respectively. 
Moreover, $^\ast F^{\mu\nu} = b^{\mu}u^{\nu} - b^{\nu}u^{\mu}$ is the (dual of the) Faraday tensor relative to an observer with 
four-velocity $u^{\mu}$, and $b^{\mu}$ is the magnetic field in that frame, with
$b^2=b^{\mu}b_{\mu}$. Assuming the magnetic field is purely azimuthal, i.e.~$b^r = b^{\theta} = 0$,
and taking into account that the flow is stationary and axisymmetric, the conservation of the current density and of the Faraday tensor follow. Contracting Eq.~\eqref{eq:e-m_tensor} with the projection tensor $h^{\alpha}_{\,\,\beta} = \delta^{\alpha}_{\,\,\beta} + u^{\alpha}u_{\beta}$, we arrive at
\begin{equation}
(w + b^2)u_{\nu}\partial_i u^{\nu} + \partial_i\left(p + \frac{b^2}{2}\right) - b_{\nu}\partial_i b^{\nu}=0\,,
\end{equation}
where $i = r, \theta$. Following~\cite{Komissarov:2006} we rewrite this equation in terms of the specific angular momentum $l$ and of the angular velocity $\Omega$, to obtain
\begin{equation}\label{eq:diff_ver}
\partial_i(\ln u_t|) - \frac{\Omega \partial_i l}{1-l\Omega} + \frac{\partial_i p}{w} + \frac{\partial_i(\mathcal{L}b^2)}{2\mathcal{L}w} = 0\,,
\end{equation}
where $\mathcal{L} = g_{t\phi}^2 - g_{tt}g_{\phi\phi}$.
To integrate Eq.~\eqref{eq:diff_ver} we first assume a barotropic equation of state $w = w(p)$ of the form
\begin{equation}\label{eq:eos_fluid}
p = K w^{\kappa},
\end{equation}
with $K$ and $\kappa$ constants.
Then, we define the magnetic pressure as $p_{\mathrm{m}} = b^2/2$, and introduce the definitions $\tilde{p}_{\mathrm{m}} = \mathcal{L} p_{\mathrm{m}}$ and $\tilde{w} = \mathcal{L} w$, in order to write an analogue equation to Eq.~\eqref{eq:eos_fluid} for $\tilde{p}_{\mathrm{m}}$~\citep{Komissarov:2006}
\begin{equation}\label{eq:eos_mag_tilde}
\tilde{p}_{\mathrm{m}} = K_{\mathrm{m}} \tilde{w}_{\mathrm{m}}^{\lambda
},
\end{equation}
or, in terms of the magnetic pressure $p_{\mathrm{m}}$
\begin{equation}\label{eq:eos_mag}
p_{\mathrm{m}} = K_{\mathrm{m}} \mathcal{L}^{\lambda
-1} w^{\lambda
},
\end{equation}
where $K_{\mathrm{m}}$ and $\lambda
$ are constants.
This particular choice of barotropic relationships, $w = w(p)$ and $\tilde{w} = \tilde{w}(\tilde{p}_{\mathrm{m}})$, fulfill the general relativistic version of the von Zeipel theorem for a toroidal magnetic field~\citep{vonZeipel:1924, Zanotti:2015}, i.e.~the surfaces of constant $\Omega$ and constant $l$ coincide.

We can now integrate Eq.~\eqref{eq:diff_ver} to obtain
\begin{equation}\label{eq:pre_full_int}
\ln |u_t| - \int^l_0 \frac{\Omega \mathrm{d}l}{1 - \Omega l} + \int^p_0 \frac{\mathrm{d}p}{w} + \int_0^{\tilde{p}_{\mathrm{m}}} \frac{\mathrm{d}\tilde{p}_{\mathrm{m}}}{\tilde{w}} = \mathrm{const}.
\end{equation}
On the surface of the disc, and particularly on its inner edge, the conditions
$p = \tilde{p}_{\mathrm{m}} = 0, \; u_t = u_{t, \mathrm{in}}, \; l = l_{\mathrm{in}}$
are satisfied and, therefore, the integration constant is simply given by
\begin{equation}
\mathrm{const.} = \ln |u_t| - \int^l_{l_\mathrm{in}} \frac{\Omega \mathrm{d}l}{1 - \Omega l}\,.
\end{equation}
We can also introduce the total (gravitational plus centrifugal) potential $W$~\citep{Abramowicz:1978} and write the integral form of the equation of motion (the relativistic Euler equation) as
\begin{equation}\label{eq:potential}
W - W_{\mathrm{in}} = \ln|u_t| - \ln|u_{t,\mathrm{in}}| - \int^{l}_{l_{\mathrm{in}}} \frac{\Omega \mathrm{d}l}{1 - \Omega l},
\end{equation}
where $W_{\mathrm{in}}$ is the potential at the inner edge of the disc (in the equatorial plane). With this definition, we can write Eq.~\eqref{eq:pre_full_int} as
\begin{equation}\label{eq:full_int}
W - W_{\mathrm{in}} = \int^p_0 \frac{\mathrm{d}p}{w} + \int_0^{\tilde{p}_{\mathrm{m}}} \frac{\mathrm{d}\tilde{p}_{\mathrm{m}}}{\tilde{w}},
\end{equation}
which for a barotropic equation of state can be easily integrated to give
\begin{equation}\label{eq:pre_enthalpy_eq}
W - W_{\rm{in}} + \frac{\kappa}{\kappa - 1}\frac{p}{w} + \frac{\lambda
}{\lambda
 - 1}\frac{p_{\mathrm{m}}}{w} = 0\,.
\end{equation}
Replacing $p$ and $p_{\mathrm{m}}$ by equations \eqref{eq:eos_fluid} and \eqref{eq:eos_mag}, the previous equation reduces to
\begin{equation}\label{eq:enthalpy_eq}
W - W_{\rm{in}} + \frac{\kappa}{\kappa - 1} K w^{\kappa - 1} + \frac{\lambda
}{\lambda
 - 1}K_{\mathrm{m}}(\mathcal{L} w)^{\lambda
 - 1} = 0,
\end{equation}
which relates the distribution of the potential with the distribution of the enthalpy density.

@@

\begin{figure*}[t]
\centering
\includegraphics[scale=0.14]{figures/fig1a.pdf}
\hspace{-0.3cm}
\includegraphics[scale=0.14]{figures/fig1b.pdf}
\hspace{-0.2cm}
\includegraphics[scale=0.14]{figures/fig1c.pdf}
\caption{Comparison with Komissarov's solution for a constant angular momentum model. The left panel shows the rest-mass density distribution in logarithm scale for all radii and all angles while the middle and right panels show, respectively, the angular profile of the logarithm of the rest-mass density at the centre of the disc and the radial profile of the logarithm of the rest-mass density at the equatorial plane.}
\label{komissarov}
\end{figure*}

Next, we choose all the initial radial values for the integration of Eq.~\eqref{eq:F} to lie between $r_{\mathrm{cusp}}$ and $r_{\mathrm{c}}$ ($\theta = \pi / 2$). Since we are only interested in the equipotential surfaces inside the Roche lobe of the disc, our choice of initial values provides us a mapping of the equipotential surfaces of the torus. Given that we have already obtained both the equipotential surfaces $\theta(r)$ which cross the segment $[r_{\mathrm{cusp}}, r_{\mathrm{c}}]$ at the equatorial plane and the values of the potential in that segment, we can obtain the complete potential distribution for the torus (outside of that segment). Once we have the potential distribution, we can find the fluid pressure at the centre of the disc from Eq.~\eqref{eq:enthalpy_eq},
\begin{equation}
p_{\mathrm{c}} = w_{\mathrm{c}}(W_{\mathrm{in}} - W_{\mathrm{c}})\left(\frac{\kappa}{\kappa - 1} + \frac{\lambda
}{\lambda
 - 1}\frac{1}{\beta_{\mathrm{m}_{\mathrm{c}}}}\right)^{-1},
\end{equation}
where $w_{\mathrm{c}}$ is the enthalpy density at the centre and
\begin{equation}
\label{eq:beta_eq}
\beta_{\mathrm{m}} = p/p_{\mathrm{m}},
\end{equation}
is the magnetisation parameter ($\beta_{\mathrm{m}_{\mathrm{c}}}$ being the magnetisation parameter at the centre of the disc). Using this definition, we can find the magnetic pressure at the centre,
\begin{equation}
p_{\mathrm{m_{\mathrm{c}}}} = p_{\mathrm{c}}/\beta_{\mathrm{m}_{\mathrm{c}}}\,.
\end{equation}
With both pressures known at the centre, fluid and magnetic, we can now find the constants $K$ and $K_{\mathrm{m}}$ using equations \eqref{eq:eos_fluid} and \eqref{eq:eos_mag}. Therefore, for a given inner radius of the disc $r_{\mathrm{in}}$ we can obtain the potential $W_{\mathrm{in}}$. We now finally have all the ingredients required to find the enthalpy density distribution (Eq.~\eqref{eq:enthalpy_eq}), the fluid pressure and magnetic pressure distributions (Eqs.~\eqref{eq:eos_fluid} and \eqref{eq:eos_mag}), and the rest-mass density distribution, which can be trivially obtained inverting the barotropic relation, $p = K \rho^{\kappa}$. We note that the fluid enthalpy density $w$ includes the rest-mass density $\rho$, and can thus be defined as $w=\rho h$ where $h$ is the specific enthalpy. The relativistic definition of this quantity is $h=1+\varepsilon+p/\rho$, where $\varepsilon$ is
the specific internal energy. From a thermodynamical point of view a non-relativistic fluid satisfies $h=1$. Therefore,
since we use poytropic equations of state (relating $p$ with either $w$ or $\rho$ in the same functional form) we are
implictly assuming that $h=1$, i.e.~the discs we build in our procedure are non-relativistic from a thermodynamical point of view.

@@

To reduce the number of free parameters to build the initial models we do as in~\citet{Komissarov:2006} and fix the values of the equation of state exponents, $\kappa$ and $\lambda$, and of the enthalpy density at the disc centre, $w_{\mathrm{c}}$. More precisely, we choose $\kappa = \lambda = 4/3$ and $w_{\mathrm{c}} = 1$. This still leaves us with five parameters to control the size, shape, thickness, and magnetisation of the disc, namely the magnetisation parameter $\beta_{\mathrm{m}_{\mathrm{c}}}$, the parameters of the angular momentum distribution $\beta$ and $\gamma$, the black hole spin parameter $a$, and the inner radius of the disc $r_{\mathrm{in}}$. We build a series of 45 models, whose main features are summarised in Table~\ref{table:1} (for $\beta_{\mathrm{m}_{\mathrm{c}}} = 10^{3}$, i.e.~models where the effects of the magnetisation are unimportant), Table~\ref{table:2} ($\beta_{\mathrm{m}_{\mathrm{c}}} = 1$, i.e.~equipartition models), and Table~\ref{table:3} ($\beta_{\mathrm{m}_{\mathrm{c}}} = 10^{-3}$, i.e.~highly-magnetised models). We note that we can build models with very small values of $\beta_{\mathrm{m}_{\mathrm{c}}}$, such as  e.g.~$\beta_{\mathrm{m}_{\mathrm{c}}} = 10^{-20}$ (extremely high magnetisation) without encountering numerical difficulties. No qualitative differences in the structure of the discs are found once $\beta_{\mathrm{m}_{\mathrm{c}}}$ becomes smaller than a sufficiently small value, about $\beta_{\mathrm{m}_{\mathrm{c}}}=10^{-3}$. This is the reason why we choose that particular value as our lower limit in the figures and tables of the manuscript.

@@

We start by assessing our procedure by first building an extra disc model that can be directly compared with one of the two models in~\citet{Komissarov:2006}. This is shown in Figure~\ref{komissarov}, which corresponds to the same constant specific angular momentum model A presented by~\citet{Komissarov:2006}. The parameters of this model are $a=0.9$, $\beta_{\mathrm{m}_{\mathrm{c}}}=0.1$, and $l=2.8$.
The visual comparison shows that our approach can reproduce those previous results with good agreement. The rest-mass density distribution in the $(r\sin\theta,r\cos\theta)$ plane shown in the left panel of Fig.~\ref{komissarov} is remarkably similar to that shown in the left panel of Fig.~2 in~\citet{Komissarov:2006}. Not only the morphology of both models is nearly identical but also the range of variation of the rest-mass density and the location of the disc centre agree well in both cases. This can be most clearly seen in the middle and right panels of Fig.~\ref{komissarov} which show, repectively, the angular profile at $r_{\rm c}$ and the radial profile at $\theta=\pi/2$. The middle figure can be directly compared with Fig.~3 of~\citet{Komissarov:2006}. It is relevant to mention that very small changes in the location of the inner radius $r_{\mathrm{in}}$ have a significant effect on the maximum value of the rest-mass density, which explains the small differences between our figures and the ones presented in~\citet{Komissarov:2006}.

@@

@@

\begin{acknowledgements}
It is a pleasure to thank P.~C.~Fragile and N.~Stergioulas for useful discussions and comments. We thank the referee, L.~Rezzolla, for his constructive report which improved the manuscript. Work supported by the Spanish MINECO (grant AYA2015-66899-C2-1-P) and the Generalitat Valenciana (PROMETEOII-2014-069).
\end{acknowledgements}
